Hadoop Image Processing Framework

### 2015-03-15 Paper.tex
Fixed spacing issue.
minor text changes and table descriptions

\subsection{Downloading and storing image data}
Hadoop uses the Hadoop Distributed File System (HDFS)\cite{Shvachko2010} to store files
in various nodes throughout the cluster.  One of Hadoop's significant
problems is that of small file storage. \cite{White2009} A small file
is one which is significantly smaller than HDFS block size.  Large
image datasets are made up of small image files in great numbers,
which is a situation which HDFS has a great deal of trouble
handling. The problem can be solved by providing a container to group
the files in some way. Hadoop offers a few options:
\begin{itemize}
	\item HAR File
	\item Sequence File
	\item Map File
\end{itemize}

@@

\textbf{Step 1: Input a URL List.} Initially users input a file
containing URLs of images to download. The input list should be a text
file with one image URL per line. The list can be generated by hand,
extracted from a database or a provided by a search. The framework
provides a extendable ripper module for extracting URLS from Flickr
and Google image searches and from SQL databases. In addition to the
list the user selects the type of image bundle to be generated
(e.g. HAR, sequence or map).  Our system divides the URLs for download
across the available processing nodes for maximum efficiency and
parallelism.  Each node map task generates several image bundles
appropriate to the selected input type, containing all of the images
it downloaded. In the reduce phase, the Reducer will merge these image
bundles into a large image bundle.

@@

\subsection{Processing image bundle using MapReduce}
Hadoop MapReduce programs handle input and output data very
efficiently, but their native data exchange formats are not convenient
for representing or manipulating image data.  For instance,
distributing images across map nodes require the translation of images
into strings, then later decoding these image strings into specified
formats in order to access pixel information.  This is both
inefficient and inconvenient. To overcome this problem, images should
be represented in as many different formats as possible, increasing
flexibility. The framework focuses on bringing familiar data types
directly to user.

As distribution is important in MapReduce, images should be processed
in the same machine where the bundle block resides. In a generic
MapReduce system, the user is responsible for creating InputFormat and
RecordReader classes to specify the MapReduce job and distribute the
input among nodes. This is a cumbersome and error-prone task; the
Hadoop Image Processing Framework provides such InputFormat and
RecordReaders for the sake of the system's ease of use.

Images are distributed as various image data types and users have
immediate access to pixel values.  If pixel values are naively
extracted from the image byte formats, valuable image header data
(e.g. JPEG EXIF data or IHDR \cite{David03} image headers) is
lost. The framework holds the image data in a special HImageHeader
data type before converting the image bytes into pixel values.  After
processing pixel data, image headers are reattached to the processed
results.  The small amount of storage overhead required for this
functionality is a trade-off worth making for the sake of preserving
image header data after processing.

@@

\textbf{Step 1: Devise the algorithm.} We assume that the user writes
an algorithm which extends the provided GenericAlgorithm class. This
class is passed as an argument to the processor module. The framework
starts a MapReduce job with the algorithm as an input. The
GenericAlgorithm holds an HImage variable; this allows user to write
an algorithm on a single image data structure, which the framework
then iterates over the entire image bundle. In addition to the
algorithm, the user should provide the image bundle file that needs to
be processed.  Depending on the specifics of the image bundle
organization and contents, the bundle is divided across nodes as
individual map tasks. Each map task will apply the processing
algorithm to each local image and append them in turnto the output
image bundle. In the reduce phase, the Reducer merges these image
bundles into a large image bundle.

@@

\textbf{Step 2: Split Image bundle across nodes} The input image
bundle is split across available nodes using the framework's custom
input format and record reader classes for maximum throughput. Once a
map task starts, HImage objects are retrieved.

\textbf{Step 3: Extract individual image} The image bytes obtained
from each HImage object are stored onto the filesystem in the
appropriate file format based on its image type.  The Extractor module
lacks a reduce phase.

@@

We chose the following algorithms: Laplacian filtering, Canny edge
detection and $k$-means image segmentation.  These are widely-used,
computation- and data-intensive algorithms of varying complexity which
require large distributed systems for timely operation on hundreds of
thousands of images.

@@

We present execution time and programming complexity results for the
three image processing algorithms described above, demonstrating the
effectiveness of the Hadoop Image Processing Framework.  These
algorithms are not particularly challenging to implement on a single
node, but producing a distributed Hadoop implementation is quite
challenging and requires a great deal of familiarity with Hadoop
concepts and distributed software engineering expertise.  On the other
hand, our framework provides all of the performance and scalability
benefits of the Hadoop system while retaining the simplicity of the
single-node implementation.  In this section we describe the dataset
and hardware and software specifications used in our experiments, and
discuss the results.

@@

\subsection{Characteristics of Image Dataset, Hardware and Software}
We performed image processing computations on a dataset of about 1TB
in size. The dataset was obtained by performing a Flickr search based
on the keyword ``flight''.  The dataset was extracted using the
FlickrRipper, an instance of the Ripper interface provided by our
Hadoop Image Processing Framework.  The framework currently provides
rippers for Flickr and Google Image searches, and more will be
provided in the future.  A Ripper interface extracts image URLs from
search results, and can be implemented for any set of results in any
format.  The experimental dataset is composed of 220000 images at 4.76
MB/image, or about 1 TB.

\subsection{Computer Hardware and Software Characterstics}
We ran the image processing computations on a 6-node distributed
cluster and on a single-node desktop computer. The table below
summarizes the cluster and desktop hardware and software
specifications. The cluster nodes differ in terms of CPU speed and
RAM. We installed Hadoop, Cloudera CDH5 and Java 1.7 on the cluster to
support Java code execution. The desktop computer had a similar
software configuration to the cluster.

\begin{table}[h]
	\begin{tabular}{|l|l|l|l|}
		\hline
		& \multicolumn{1}{c|}{\textbf{Specs}}                            & \multicolumn{1}{c|}{\textbf{Cluster}}                                                                                                           & \multicolumn{1}{c|}{\textbf{Desktop}}                                                                               \\ \hline
		\multirow{2}{*}{Hardware} & Cluster Nodes                                                  & \begin{tabular}[c]{@{}l@{}}6 data nodes with \\ from 2 to 4 virtual \\ processors and 4 GB\\ RAM and name node \\ with 8 GB RAM\end{tabular} & \begin{tabular}[c]{@{}l@{}}Intel Xeon @ 2Ghz\\ 8 cores, 16GB of RAM \\ and hyperthreading\\ activated\end{tabular} \\ \cline{2-4} 
		& Networking                                                     & 1 Gbit/second                                                                                                                                    &                                                                                                                     \\ \hline
		\multirow{3}{*}{Software} & \begin{tabular}[c]{@{}l@{}}Java Virtual\\ Machine\end{tabular} & \begin{tabular}[c]{@{}l@{}}Java version \\ "1.7.0\_45"\\ Java SE Runtime\\ Environment\end{tabular}                                             & \begin{tabular}[c]{@{}l@{}}Java version \\ "1.7.0\_45"\\ Java SE Runtime\\ Environment\end{tabular}                 \\ \cline{2-4} 
		& Hadoop                                                         & \begin{tabular}[c]{@{}l@{}}Hadoop 2.5\\ Cloudera CDH5\end{tabular}                                                                              &                                                                                                                     \\ \cline{2-4} 
		& \begin{tabular}[c]{@{}l@{}}Operating\\ System\end{tabular}     & CentOS 6.5                                                                                                                                      & CentOS 6.5                                                                                                          \\ \hline
	\end{tabular}
\end{table}

@@

Each algorithm is also implemented on a Hadoop cluster without using
the Hadoop Image Processing Framework.  These implementations overcome
the problem of handling large datasets but they suffer from extremely
complex code.  The user needs to write a great deal of code and
understand the technical details of Hadoop. Writing such code,
including custom InputFormats and RecordReaders, is cumbersome and
error-prone. Fig. \ref{fig:module-chart} compares the actual lines of
code a user needs to write in order to run our experimental algorithms
(or any similar image processing application) on a Hadoop cluster.

@@

\begin{table}[h]
	\begin{tabular}{|l|l|l|l|}
		\hline
		\multicolumn{1}{|c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Computa-\\ tional\\ Platform\end{tabular}}} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Computational \\ Elasticity\end{tabular}}}      & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Code \\ Complexity\end{tabular}}}                                                                                                                                                  & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Data \\ Locality\end{tabular}}}                                                                           \\ \hline
		Desktop                                                                                             & \begin{tabular}[c]{@{}l@{}}Low : limited by \\ RAM and CPU of \\ the executing\\ computer\end{tabular} & \begin{tabular}[c]{@{}l@{}}Normal : Complexity \\ lies in writing the \\ algorithm\end{tabular}                                                                                                                                           & \begin{tabular}[c]{@{}l@{}}All data are \\ on local disk\end{tabular}                                                                                            \\ \hline
		\begin{tabular}[c]{@{}l@{}}Hadoop:\\ without \\ using \\ framework\end{tabular}                     & \begin{tabular}[c]{@{}l@{}}High: Nodes can\\  be requested \\ as needed\end{tabular}                   & \begin{tabular}[c]{@{}l@{}}Heavy : Complexity \\ lies in writing every \\ module in framework \\ including custom \\ InputFormats and \\ RecordReaders.  User \\ requires high \\ technical expertise \\ with Hadoop.\end{tabular} & \begin{tabular}[c]{@{}l@{}}All data resides\\ in HDFS. Custom\\ InputFormats \\ need to be \\ created to launch\\ computations \\ in appropriate \\ locations\end{tabular} \\ \hline
		\begin{tabular}[c]{@{}l@{}}Hadoop:\\ using \\ framework\end{tabular}                                & \begin{tabular}[c]{@{}l@{}}High: Nodes can \\ be requested \\ as needed\end{tabular}                   & \begin{tabular}[c]{@{}l@{}}Normal : Complexity \\ lies in writing \\ algorithm and is \\ equivalent to writing \\ on desktop platform\end{tabular}                                                                                        & \begin{tabular}[c]{@{}l@{}}All data resides \\ in HDFS. Highly \\ parallelized.\end{tabular}                                                                     \\ \hline
	\end{tabular}
\end{table}

@@

\begin{itemize}
\item Three image processing algorithms described in Section
  \ref{algorithms}
\item Datasets of varying image counts
\item Configurations of hardware cluster used for computations 
\item Coding complexity (measured in Lines of Code (LOC)
\end{itemize}

@@

The Hadoop Image processing framework provides transparency on
multiple levels.  Users can use the predefined MapReduce modules
(Downloader, Processor and Extractor) for processing the image
modules, and most image processing tasks require no more than this.
Users may also write custom MapReduce tasks within the framework,
taking advantage of its hierarchical construction.  This still
protects the user from the full complexity and error-prone nature of
the full Hadoop framework, but requires additional knowledge and
expertise on the user's part.

@@

\begin{lstlisting}[caption = Comparing File Writer instance in java and SequenceBundleWriter instance in Hadoop Image Processing framework ]

//Sample File Writing in Java

File file = new File("abc.txt");
FileWriter fw=new FileWriter(file);
fw.append(val)
fw.close();

//Sample BundleFile Writing in Hadoop using Framework

BundleFile file = new BundleFile("abc.seq");
SequenceBundleWriter sbw=new SequenceBundleWriter(file);
sbw.append(himage);
sbw.close();

\end{lstlisting}

>>>>

\subsection{Downloading and storing image data}
Hadoop uses the Hadoop Distributed File System (HDFS)\cite{Shvachko2010} to store files
in various nodes throughout the cluster.  One of Hadoop's significant
problems is that of small file storage. \cite{White2009} A small file
is one which is significantly smaller than HDFS block size.  Large
image datasets are made up of small image files in great numbers,
which is a situation HDFS has a great deal of trouble
handling. This problem can be solved by providing a container to group
the files in some way. Hadoop offers a few options:
\begin{itemize}
	\item HAR File
	\item Sequence File
	\item Map File
\end{itemize}

@@

\textbf{Step 1: Input a URL List.} Initially users input a file
containing URLs of images to download. The input list should be a text
file with one image URL per line. The list can be generated by hand,
extracted from a database or a provided by a search. The framework
provides an extendable ripper module for extracting URLs from Flickr
and Google image searches and from SQL databases. In addition to the
list the user selects the type of image bundle to be generated
(e.g. HAR, sequence or map).  Our system divides the URLs for download
across the available processing nodes for maximum efficiency and
parallelism. The URL list is split into several map tasks of equal size across the nodes. Each node map task generates several image bundles
appropriate to the selected input list, containing all of the image URLs
to download. In the reduce phase, the Reducer will merge these image
bundles into a large image bundle.

@@

\subsection{Processing image bundle using MapReduce}
Hadoop MapReduce program handles input and output data very
efficiently, but their native data exchange formats are not convenient
for representing or manipulating image data.  For instance,
distributing images across map nodes require the translation of images
into strings, then later decoding these image strings into specified
formats in order to access pixel information.  This is both
inefficient and inconvenient. To overcome this problem, images should
be represented in as many different formats as possible, increasing
flexibility. The framework focuses on bringing familiar data types
directly to user.

As distribution is important in MapReduce, images should be processed
in the same machine where the bundle block resides. In a generic
MapReduce system, the user is responsible for creating InputFormat and
RecordReader classes to specify the MapReduce job and distribute the
input among nodes. This is a cumbersome and error-prone task; the
Hadoop Image Processing Framework provides such InputFormat and
RecordReaders for system's ease of use.

Images are distributed as various image data types and users have
immediate access to pixel values.  If pixel values are naively
extracted from the image byte formats, valuable image header data
(e.g. JPEG EXIF data or IHDR \cite{David03} image headers) is
lost. The framework holds the image data in a special HImageHeader
data type before converting the image bytes into pixel values.  After
processing pixel data, image headers are reattached to the processed
results.  The small amount of storage overhead required for this
functionality is a trade-off worth making for preserving
image header data after processing.

@@

\textbf{Step 1: Devise the algorithm.} We assume that the user writes
an algorithm which extends the provided GenericAlgorithm class. This
class is passed as an argument to the processor module. The framework
starts a MapReduce job with the algorithm as an input. The
GenericAlgorithm holds an HImage variable; this allows user to write
an algorithm on a single image data structure, which the framework
then iterates over the entire image bundle. In addition to the
algorithm, the user should provide the image bundle file that needs to
be processed.  Depending on the specifics of the image bundle
organization and contents, the bundle is divided across nodes as
individual map tasks. Each map task will apply the processing
algorithm to each local image and append them to the output
image bundle. In the reduce phase, the Reducer merges these image
bundles into a large image bundle.

@@

\textbf{Step 2: Split Image bundle across nodes.} The input image
bundle is split across available nodes using the framework's custom
input format and record reader classes for maximum throughput. Once a
map task starts, HImage objects are retrieved.

\textbf{Step 3: Extract individual image.} The image bytes obtained
from each HImage object are stored onto the filesystem in the
appropriate file format based on its image type.  The Extractor module
lacks a reduce phase.

@@

We chose the following algorithms: Laplacian filtering, Canny edge
detection and $k$-means image segmentation.  These are widely-used,
computation-intensive and data-intensive algorithms of varying complexity which
require large distributed systems for timely operation on hundreds of
thousands of images.

@@

We present execution time and programming complexity results for the
three image processing algorithms described above, demonstrating the
effectiveness of the Hadoop Image Processing Framework.  These
algorithms are not particularly challenging to implement on a single
node, but producing a distributed Hadoop implementation is quite
challenging and requires a great deal of familiarity with Hadoop
concepts and distributed software engineering expertise.  On the other
hand, our framework provides all of the performance and scalability
benefits of the Hadoop system while retaining the simplicity of the
single-node implementation.  In this section we describe the dataset, 
hardware and software specifications used in our experiments, and
discuss the results.

@@

\subsection{Characteristics of Image Dataset, Hardware and Software}
We performed image processing computations on a dataset of about 1 TB
in size. The dataset was obtained by performing a Flickr search based
on the keyword ``flight''.  The dataset was extracted using the
FlickrRipper, an instance of the Ripper interface provided by our
Hadoop Image Processing Framework.  The framework currently provides
rippers for Flickr and Google image searches, and more will be
provided in the future.  A Ripper interface extracts image URLs from
search results, and can be implemented for any set of results in any
format.  The experimental dataset is composed of 220,000 images at 4.76
MB/image, or about 1 TB.

\subsection{Computer hardware and software characteristics}
We ran the image processing computations on a 6-node distributed
cluster and on a single-node desktop computer. The table below
summarizes the cluster and desktop hardware and software
specifications. The cluster nodes differ in terms of CPU speed and
RAM. We installed Hadoop, Cloudera CDH5 and Java 1.7 on the cluster to
support Java code execution. The desktop computer had a similar
software configuration to the cluster.

\begin{table}[h]
	\caption {Computer hardware and software characteristics.}
	\label{table:table1}
	\begin{tabular}{|l|l|l|l|}
		\hline
		& \multicolumn{1}{c|}{\textbf{Specs}}                            & \multicolumn{1}{c|}{\textbf{Cluster}}                                                                                                           & \multicolumn{1}{c|}{\textbf{Desktop}}                                                                               \\ \hline
		\multirow{2}{*}{Hardware} & Cluster Nodes                                                  & \begin{tabular}[c]{@{}l@{}}6 data nodes with \\ from 2 to 4 virtual \\ processors and 4 GB\\ RAM and name node \\ with 8 GB RAM\end{tabular} & \begin{tabular}[c]{@{}l@{}}Intel Xeon @ 2Ghz\\ 8 cores, 16GB of RAM \\ and hyperthreading\\ activated\end{tabular} \\ \cline{2-4} 
		& Networking                                                     & 1 Gbit/second                                                                                                                                    &                                                                                                                     \\ \hline
		\multirow{3}{*}{Software} & \begin{tabular}[c]{@{}l@{}}Java Virtual\\ Machine\end{tabular} & \begin{tabular}[c]{@{}l@{}}Java version \\ "1.7.0\_45"\\ Java SE Runtime\\ Environment\end{tabular}                                             & \begin{tabular}[c]{@{}l@{}}Java version \\ "1.7.0\_45"\\ Java SE Runtime\\ Environment\end{tabular}                 \\ \cline{2-4} 
		& Hadoop                                                         & \begin{tabular}[c]{@{}l@{}}Hadoop 2.5\\ Cloudera CDH5\end{tabular}                                                                              &                                                                                                                     \\ \cline{2-4} 
		& \begin{tabular}[c]{@{}l@{}}Operating\\ System\end{tabular}     & CentOS 6.5                                                                                                                                      & CentOS 6.5                                                                                                          \\ \hline
	\end{tabular}

\end{table}

@@

Each algorithm is also implemented on a Hadoop cluster without using
the Hadoop Image Processing Framework.  These implementations overcome
the problem of handling large datasets but they suffer from extremely
complex code.  The user needs to write a great deal of code and
understand the technical details of Hadoop. Writing such code,
including custom InputFormats and RecordReaders, is cumbersome and
error-prone. Figure \ref{fig:module-chart} compares the actual lines of
code a user needs to write in order to run our experimental algorithms
(or any similar image processing application) on a Hadoop cluster.

@@

\begin{table}[h]
	\caption{Characteristics of computational elasticity, coding complexity and data locality for the three experimental platforms.}
	\label{table:table2}
	\begin{tabular}{|l|l|l|l|}
		\hline
		\multicolumn{1}{|c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Computa-\\ tional\\ Platform\end{tabular}}} & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Computational \\ Elasticity\end{tabular}}}      & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Code \\ Complexity\end{tabular}}}                                                                                                                                                  & \multicolumn{1}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Data \\ Locality\end{tabular}}}                                                                           \\ \hline
		Desktop                                                                                             & \begin{tabular}[c]{@{}l@{}}Low : limited by \\ RAM and CPU of \\ the executing\\ computer\end{tabular} & \begin{tabular}[c]{@{}l@{}}Normal : Complexity \\ lies in writing the \\ algorithm\end{tabular}                                                                                                                                           & \begin{tabular}[c]{@{}l@{}}All data are \\ on local disk\end{tabular}                                                                                            \\ \hline
		\begin{tabular}[c]{@{}l@{}}Hadoop:\\ without \\ using \\ framework\end{tabular}                     & \begin{tabular}[c]{@{}l@{}}High: Nodes can\\  be requested \\ as needed\end{tabular}                   & \begin{tabular}[c]{@{}l@{}}Heavy : Complexity \\ lies in writing every \\ module in framework \\ including custom \\ InputFormats and \\ RecordReaders.  User \\ requires high \\ technical expertise \\ with Hadoop.\end{tabular} & \begin{tabular}[c]{@{}l@{}}All data resides\\ in HDFS. Custom\\ InputFormats \\ need to be \\ created to launch\\ computations \\ in appropriate \\ locations\end{tabular} \\ \hline
		\begin{tabular}[c]{@{}l@{}}Hadoop:\\ using \\ framework\end{tabular}                                & \begin{tabular}[c]{@{}l@{}}High: Nodes can \\ be requested \\ as needed\end{tabular}                   & \begin{tabular}[c]{@{}l@{}}Normal : Complexity \\ lies in writing \\ algorithm and is \\ equivalent to writing \\ on desktop platform\end{tabular}                                                                                        & \begin{tabular}[c]{@{}l@{}}All data resides \\ in HDFS. Highly \\ parallelized.\end{tabular}                                                                     \\ \hline
	\end{tabular}
\end{table}

@@

\begin{itemize}
\item Three image processing algorithms described in Section
  \ref{algorithms}
\item Datasets of varying image counts
\item Configurations of hardware cluster used for computations 
\item Coding complexity - measured in Lines of Code (LOC)
\end{itemize}

@@

The Hadoop Image Processing Framework provides transparency on
multiple levels.  Users can use the predefined MapReduce modules
(Downloader, Processor and Extractor) for processing the image
modules, and most image processing tasks require no more than this.
Users may also write custom MapReduce tasks within the framework,
taking advantage of its hierarchical construction.  This still
protects the user from the full complexity and error-prone nature of
the full Hadoop framework, but requires additional knowledge and
expertise on the user's part.

@@

\begin{lstlisting}[caption = Comparing FileWriter instance in java and SequenceBundleWriter instance in Hadoop Image Processing framework ]
//Sample File Writing in Java

File file = new File("test");
FileWriter fw = new FileWriter(file);
fw.append(val)
fw.close();
//Sample BundleFile Writing in Hadoop using Framework

BundleFile file = new BundleFile("test_bundle.seq");
SequenceBundleWriter sbw = new SequenceBundleWriter(file);
sbw.append(himage);
sbw.close();
\end{lstlisting}
